Heuristics for the Score-Constrained Strip-Packing Problem

### 2018-06-07 PAPER/main.tex
added n = |I|

A prominent application of the COP is in a strip-packing problem brought to light as an open-combinatorial problem by \cite{goulimis2004}. Here, a set $\mathcal{I}$ of rectangular items of equal height $H$ made from cardboard are to be packed onto a strip of height $H$ from left to right. Each item $i \in \mathcal{I}$ has width $w_i \in \mathbb{Z}^{+}$, and possesses two vertical score lines, marked in predetermined places. A pair of knives mounted onto a bar cuts along the score lines of two adjacent items simultaneously, which allows the items to be folded with ease (see Figure \ref{fig:itemsknife}). However, by design, the scoring knives cannot be placed too close to one another, and as such have a ``minimum scoring distance'' (around 70mm in industry). The distances between each score line and the nearest edge on an item $i \in \mathcal{I}$ are the score widths, $a_i, b_i \in \mathbb{Z}^{+}$ where $a_i + b_i < w_i$, assigned such that $a_i \leq b_i$. Since these score widths are not necessarily equal, an item $i$ can be positioned in one of two orientations: ``regular'', denoted $(a_i, b_i)$, or ``rotated'', denoted $(b_i, a_i)$, where the smaller of the two score widths $a_i$ is on the left- and right- hand side of the item respectively.

@@

\begin{definition}
	Let $\mathcal{I}$ be a set of rectangular items of height $H$ with varying widths $w_i \in \mathbb{Z}^+$ and score widths $a_i, b_i \in \mathbb{Z}^+$ for each item $i \in \mathcal{I}$. Given a minimum scoring distance $\tau \in \mathbb{Z}^+$, the Score-Constrained Strip-Packing Problem (SCSPP) consist of finding the minimum number of strips of height $H$ and width $W$ required to pack all items in $\mathcal{I}$ such that the sum of every pair of adjacent score widths is greater than or equal to $\tau$, and that no strip is overfilled.
\end{definition}

@@

An optimal solution $\mathcal{S}$ for the SCSPP is a solution that consists of the fewest number of strips $k$ needed to feasibly contain the $|\mathcal{I}|$ items in the given problem instance. A simple lower bound for $k$ is the theoretical minimum
\begin{equation}
\label{eqn:lowerbound}
t = \ceil*{\frac{\sum_{i=1}^{|\mathcal{I}|}w_i}{W}},
\end{equation}
which can be computed in linear time (\citealp{martello1990b}). Perhaps the simplest and most well-known heuristic for packing problems is \textsc{First-Fit} (FF), a greedy online algorithm that places each item presented in some arbitrary order onto the lowest-indexed strip such that the capacity of the strip is not exceeded. It is known that there always exists at least one ordering of the items such that FF produces an optimal solution (\citealp{lewis2009}). A minor modification to FF yields the \textsc{First-Fit Decreasing} (FFD) heuristic, which sorts the items in non-increasing order of size prior to performing FF. In 1973, \citeauthor{johnson1973} showed that FFD is guaranteed to return a solution that uses no more than $\frac{11}{9}k + 4$ strips. More recently, \cite{dosa2007} has proven that the worst case for FFD is in fact $\frac{11}{9}k + \frac{6}{9}$, and that this bound is tight. Due to the initial sorting of the items in non-increasing order of sizes, the time complexity of FFD is $O(n\log n)$, where $n = |\mathcal{I}|$ is the input size.

As mentioned in the introduction, the SCSPP shares many similarities with the BPP, however the addition of constraint (\ref{eqn:vscstrip}) induces complications that must be taken into consideration. One obvious difference is the order in which the items must appear on the strips. The order of the items in the BPP is unimportant, as there are no conditions regarding which items can be placed together, however in the SCSPP the items must be ordered in a way that meets the vicinal sum constraint. In addition, removing an item in the BPP retains feasibility, whereas in the SCSPP this is not guaranteed, as it may leave a subset of items for which the vicinal sum constraint is not satisfied. Furthermore, the theoretical minimum $t$ (Equation (\ref{eqn:lowerbound})) has the potential to be less accurate for the SCSPP, as the minimum scoring distance $\tau$ is not considered. For example, if the minimum scoring distance is greater than twice the largest score width, then it is clear that $|\mathcal{I}|$ strips will be required, regardless of the items' widths.
% Time complexity for FFD also inaccurate

@@

\section{Comparison}
For our experiments, we generated 1000 problem instances for three different sizes of $\mathcal{I}$, containing 100, 500, and 1000 items. For each problem instance, the items have varying widths $w_i$ between 150 and 1000, and score widths $a_i, b_i$ between 1 and 70, all selected randomly from a uniform distribution. Three strip widths $W$ were used, 5000, 2500, and 1250. As the width of the strips decrease, the average number of items per strip also decreases, making the problem more difficult. Both the items and the strips have equal height of $H = 1$. Similarly to experiments performed by \cite{lewis2011}, we also introduced a parameter $\delta$ to denote the proportion of pairs of score widths from different items that meet the vicinal sum constraint, i.e. whose sum is greater than or equal to $\tau$. Values of $\delta$ from 0.0 to 1.0 were created by changing the value of the minimum scoring distance $\tau$. Clearly, when $\delta = 0.0$, there are no score widths that can be packed together feasibly, and so $|\mathcal{I}|$ strips will be required (one for each item), whereas if $\delta = 1.0$ then all pairs of score widths meet the constraint, and the problem is analogous to the BPP.

The heuristics were implemented in C++ and executed on a computer equipped with an Intel Core i3-2120 processor with a frequency of 3.30GHz. The tables below show the average optimality of the heuristics for each value of $\delta$ from 1000 instances, where the optimality of each instance is the ratio of the number of strips in the final solution obtained by the heuristic to the theoretical minimum $t$, which was calculated using Equation \ref{eqn:lowerbound}.

Firstly, it is apparent that the heuristics perform comparitively in the same manner regardless of the input size.

\begin{itemize}
	\item \textbf{PairSmallest:}
	\item Best average optimality over all three set sizes for $W = 5000$ and $W = 2500$ for $\delta = 0.1 - 0.6$, in some cases up to $\delta = 0.8$.
	\item However, when $W = 1250$, PS only has best average optimality for $\delta = 0.1$ and $\delta = 0.2$. 
	\item PS is able to obtain solution that are on average within 10\% of the lower bound for the optimal solution for $\delta = 0.6$ onwards when $W=5000$ and $W = 2500$.
	\item When $W = 1250$, PS only obtains solutions within 10\% of the lowerbound for $\delta = 0.8$ onwards with set sizes 500 and 1000. For $|\mathcal{I}| = 100$, PS is not able to (on average) obtain solutions within 10\% of the lower bound for any value of $\delta$.
	\item Clear pattern across all three set sizes, PS performs better when the strip widths are wider, however PS is still the best of the three heuristics for smaller strip widths and harder problems, i.e. when $\delta$ is smaller. 
	\item PS never obtains avg solution within 1\% of lower bound.
\end{itemize}

\begin{itemize}
	\item optimality converge toward 1 as $\delta$ increases, i.e. as the problem instances become less difficult.
	\item When $\delta = 1.0$. MFFD and MFFD$^+$ perform identically, as the problems are equivalent to the original BPP, and so they behave like the original FFD heuristic.
	\item When $\delta = 0.0$, all optimalities are equal, as each item requires its own strip, and so $|\mathcal{I}|$ strips are needed for all 1000 instances.
	\item Values in bold represent the best avg optimality of the three heuristics for the given level of $\delta$. 
	\item Clear pattern across both set sizes, PS performs the best out of the three for $\delta = 0.1$ to 0.7 when $W = 5000$, and for $\delta = 0.1$ to 0.6 when $W = 2500$.
	\item However, when $W = 1250$, PS is only superior when $\delta = 0.1$ and 0.2. 
	\item On further inspection, we can see that the difference between the average optimalities for PS and MFFD$^+$ is negligible.
	\item Using Table 1 as an example, when $\delta = 0.2$, the solutions generated by PS are on average within 31.781\% of $t$, and for MFFD$^+$ within 32.61\%, a difference of 0.829\%. 
	\item In practical terms, PS produces solutions using 303.95 strips on average, where $t = 230.648$. MFFD$^+$ produces solutions using 305.862 strips, a difference of less than two strips. 
	\item Another trend is that PS produces average solutions within 10\% of $t$ for $\delta = 0.5$ onwards when $W = 5000$ and 2500, however when $W = 1250$, PS only produces solutions within 10\% of $t$ for $\delta = 0.8$ onwards.
	\item In contrast, MFFD and MFFD$^+$ can produce solution within 10\% of $t$ when $W = 1250$ for $\delta = 0.5$ onwards for $n = 500$, and $\delta = 0.4$ for $n = 1000$.
	\item MFFD$^+$ can also find solutions within 1\% of $t$ for $\delta = 0.8$ onwards for all strip widths in both set sizes.
	\item Number of items per strip smaller, problem is harder, falkenauer 1996, martello and toth 1990b items per strip roughly 2
	\item all instances run in less than 160ms
	\item For each instance where opt $>1$, two cases: either $t$ is incorrect, and there does not exist a solution with $k = t$ strips, and the solution found is actually an optimal solution, or $t$ is correct, and the heuristic cannot find a solution with $t$ strips, not an optimal solution.
	\item since this is a relatively new problem, no problem instances solved to optimality, and so we have no way of knowing if in fact the heuristics are finding optimal solutions. Only able to compare with the theoretical minimum. 
\end{itemize}

\begin{table}[!htb]
	\centering
	\ra{1.3}
	\caption{Average optimality ratios for $|\mathcal{I}| = 500$.}
	\begin{tabular}{@{}c@{\hspace{25pt}}c@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}c@{}}\toprule
		& \multicolumn{3}{c}{$W = 5000$, $t = 58.039$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 2500$, $t = 115.571$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 1250$, $t = 230.648$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0 & 8.61813 & 8.61813 & 8.61813 && 4.32789 & 4.32789 & 4.32789 && 2.16858 & 2.16858 & 2.16858 \\
		0.1 & 5.21367 & \textbf{4.84177} & 5.16085 && 2.65934 & \textbf{2.47735} & 2.65679 && 1.51486 & \textbf{1.47896} & 1.51486 \\
		0.2 & 4.03055 & \textbf{3.45946} & 3.97645 && 2.12099 & \textbf{1.84737} & 2.11788 && 1.32613 & \textbf{1.31781} & 1.32610 \\
		0.3 & 3.11091 & \textbf{2.34803} & 3.03816 && 1.72955 & \textbf{1.39676} & 1.70772 && 1.19451 & 1.22922 & \textbf{1.19381} \\
		0.4 & 2.43605 & \textbf{1.52907} & 2.29713 && 1.45976 & \textbf{1.12798} & 1.40951 && 1.11037 & 1.18065 & \textbf{1.10819} \\
		0.5 & 1.91057 & \textbf{1.04130} & 1.69129 && 1.26343 & \textbf{1.03341} & 1.19575 && 1.05771 & 1.15415 & \textbf{1.05319} \\
		0.6 & 1.49086 & \textbf{1.01263} & 1.24645 && 1.12448 & \textbf{1.02957} & 1.06874 && 1.02932 & 1.13470 & \textbf{1.02398} \\
		0.7 & 1.19571 & \textbf{1.01257} & 1.04468 && 1.04933 & 1.02774 & \textbf{1.01918} && 1.01741 & 1.11368 & \textbf{1.01444} \\
		0.8 & 1.04977 & 1.01203 & \textbf{1.00815} && 1.01595 & 1.02661 & \textbf{1.00790} && 1.01297 & 1.09136 & \textbf{1.01197} \\
		0.9 & 1.00929 & 1.01197 & \textbf{1.00463} && 1.00734 & 1.02586 & \textbf{1.00641} && 1.01159 & 1.07283 & \textbf{1.01139} \\
		1.0 & \textbf{1.00423} & 1.01171 & \textbf{1.00423} && \textbf{1.00626} & 1.02561 & \textbf{1.00626} && \textbf{1.01130} & 1.06549 & \textbf{1.01130} \\
		\bottomrule
	\end{tabular}	
	\label{table:n=500}
\end{table}

\begin{table}[!htb]
	\centering
	\ra{1.3}
	\caption{Average optimality ratios for $|\mathcal{I}| = 1000$.}
	\begin{tabular}{@{}c@{\hspace{25pt}}c@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}c@{}}\toprule
		& \multicolumn{3}{c}{$W = 5000$, $t = 115.534$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 2500$, $t = 230.563$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 1250$, $t = 460.623$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0 & 8.65711 & 8.65711 & 8.65711 && 4.33803 & 4.33803 & 4.33803 && 2.17138 & 2.17138 & 2.17138 \\
		0.1 & 5.17263 & \textbf{4.84248} & 5.14027 && 2.63606 & \textbf{2.48111} & 2.64271 && 1.51056 & \textbf{1.46744} & 1.51121 \\
		0.2 & 3.97565 & \textbf{3.46163} & 3.95164 && 2.09301 & \textbf{1.85824} & 2.10161 && 1.31785 & \textbf{1.31090} & 1.31876 \\
		0.3 & 3.04866 & \textbf{2.35023} & 3.01260 && 1.69777 & \textbf{1.40862} & 1.68776 && \textbf{1.18304} & 1.22111 & 1.18344 \\
		0.4 & 2.37436 & \textbf{1.52046} & 2.26560 && 1.42646 & \textbf{1.13101} & 1.38430 && 1.09722 & 1.17063 & \textbf{1.09614} \\
		0.5 & 1.84748 & \textbf{1.02600} & 1.64169 && 1.22960 & \textbf{1.03017} & 1.16950 && 1.04411 & 1.14430 & \textbf{1.03986} \\
		0.6 & 1.43331 & \textbf{1.01212} & 1.20264 && 1.09935 & \textbf{1.02738} & 1.05010 && 1.01971 & 1.12392 & \textbf{1.01480} \\
		0.7 & 1.15500 & \textbf{1.01198} & 1.02574 && 1.03421 & 1.02569 & \textbf{1.01224} && 1.01150 & 1.10164 & \textbf{1.00926} \\
		0.8 & 1.03182 & 1.01182 & \textbf{1.00587} && 1.01111 & 1.02454 & \textbf{1.00682} && 1.00861 & 1.07791 & \textbf{1.00804} \\
		0.9 & 1.00707 & 1.01165 & \textbf{1.00419} && 1.00663 & 1.02406 & \textbf{1.00608} && 1.00784 & 1.05885 & \textbf{1.00774} \\
		1.0 & \textbf{1.00399} & 1.01169 & \textbf{1.00399} && \textbf{1.00603} & 1.02400 & \textbf{1.00603} && \textbf{1.00770} & 1.05406 & \textbf{1.00770} \\
		\bottomrule
	\end{tabular}
	\label{table:n=1000}	
\end{table}

>>>>

A prominent application of the COP is in a strip-packing problem brought to light as an open-combinatorial problem by \cite{goulimis2004}. Here, a set $\mathcal{I}$ of rectangular items (where $|\mathcal{I}| = n$) of equal height $H$ made from cardboard are to be packed onto a strip of height $H$ from left to right. Each item $i \in \mathcal{I}$ has width $w_i \in \mathbb{Z}^{+}$, and possesses two vertical score lines, marked in predetermined places. A pair of knives mounted onto a bar cuts along the score lines of two adjacent items simultaneously, which allows the items to be folded with ease (see Figure \ref{fig:itemsknife}). However, by design, the scoring knives cannot be placed too close to one another, and as such have a ``minimum scoring distance'' (around 70mm in industry). The distances between each score line and the nearest edge on an item $i \in \mathcal{I}$ are the score widths, $a_i, b_i \in \mathbb{Z}^{+}$ where $a_i + b_i < w_i$, assigned such that $a_i \leq b_i$. Since these score widths are not necessarily equal, an item $i$ can be positioned in one of two orientations: ``regular'', denoted $(a_i, b_i)$, or ``rotated'', denoted $(b_i, a_i)$, where the smaller of the two score widths $a_i$ is on the left- and right- hand side of the item respectively.

@@

\begin{definition}
	Let $\mathcal{I}$ be a set of rectangular items of height $H$ with varying widths $w_i \in \mathbb{Z}^+$ and score widths $a_i, b_i \in \mathbb{Z}^+$ for each item $i \in \mathcal{I}$. Given a minimum scoring distance $\tau \in \mathbb{Z}^+$, the Score-Constrained Strip-Packing Problem (SCSPP) consists of finding the minimum number of strips of height $H$ and width $W$ required to pack all $n$ items in $\mathcal{I}$ such that the sum of every pair of adjacent score widths is greater than or equal to $\tau$, and that no strip is overfilled.
\end{definition}

@@

An optimal solution $\mathcal{S}$ for the SCSPP is a solution that consists of the fewest number of strips $k$ needed to feasibly contain the $n$ items in the given problem instance. A simple lower bound for $k$ is the theoretical minimum
\begin{equation}
\label{eqn:lowerbound}
t = \ceil*{\frac{\sum_{i=1}^{n}w_i}{W}},
\end{equation}
which can be computed in $O(n)$ time (\citealp{martello1990b}). Perhaps the simplest and most well-known heuristic for packing problems is \textsc{First-Fit} (FF), a greedy online algorithm that places each item presented in some arbitrary order onto the lowest-indexed strip such that the capacity of the strip is not exceeded. It is known that there always exists at least one ordering of the items such that FF produces an optimal solution (\citealp{lewis2009}). A minor modification to FF yields the \textsc{First-Fit Decreasing} (FFD) heuristic, which sorts the items in non-increasing order of size prior to performing FF. In 1973, \citeauthor{johnson1973} showed that FFD is guaranteed to return a solution that uses no more than $\frac{11}{9}k + 4$ strips. More recently, \cite{dosa2007} has proven that the worst case for FFD is in fact $\frac{11}{9}k + \frac{6}{9}$, and that this bound is tight. Due to the initial sorting of the items in non-increasing order of sizes, the time complexity of FFD is $O(n\log n)$.

As mentioned in the introduction, the SCSPP shares many similarities with the BPP, however the addition of constraint (\ref{eqn:vscstrip}) induces complications that must be taken into consideration. One obvious difference is the order in which the items must appear on the strips. The order of the items in the BPP is unimportant, as there are no conditions regarding which items can be placed together, however in the SCSPP the items must be ordered in a way that meets the vicinal sum constraint. In addition, removing an item in the BPP retains feasibility, whereas in the SCSPP this is not guaranteed, as it may leave a subset of items for which the vicinal sum constraint is not satisfied. Furthermore, the theoretical minimum $t$ (Equation (\ref{eqn:lowerbound})) has the potential to be less accurate for the SCSPP, as the minimum scoring distance $\tau$ is not considered. For example, if the minimum scoring distance is greater than twice the largest score width, then it is clear that $n$ strips will be required, regardless of the items' widths.
% Time complexity for FFD also inaccurate

@@

\section{Comparison}
For our experiments, we generated 1000 problem instances for two different sizes of $\mathcal{I}$ containing 500 and 1000 items. For each problem instance the items have varying widths $w_i$ between 150 and 1000 and score widths $a_i, b_i$ between 1 and 70 to ensure that each item has exactly two score lines, all selected randomly from a uniform distribution. Strips of widths $W = 5000$, 2500 and 1250 were used to alter the number of items per strip. Both the items and the strips have equal height of $H = 1$. Similarly to experiments performed by \cite{lewis2011}, we also introduced a parameter $\delta$ to denote the proportion of pairs of score widths from different items that meet the vicinal sum constraint, i.e. whose sum is greater than or equal to $\tau$. Values of $\delta$ from 0.0 to 1.0 were created by changing the value of the minimum scoring distance $\tau$. Clearly, when $\delta = 0.0$, there are no score widths that can be packed together feasibly, and so $n$ strips will be required (one for each item), whereas if $\delta = 1.0$ all pairs of score widths meet the constraint, and the problem is analogous to the BPP. (As the width of the strips decrease, the average number of items per strip also decreases, making the problem more difficult.)

The heuristics were implemented in C++ and executed on a computer equipped with an Intel Core i3-2120 processor with a frequency of 3.30GHz. We calculated the average optimality of the heuristics for each value of $\delta$ from 1000 instances, where the optimality of each instance is the ratio of the number of strips in the final solution obtained by the heuristic to the theoretical minimum $t$, which was calculated using Equation \ref{eqn:lowerbound}. All individual trials were executed in under 160ms.

Tables \ref{table:n=500} and \ref{table:n=1000} compare the results obtained from the three heuristics using different values of $\delta$ and $W$ for $n = 500$ and 1000 respectively. Figures in bold indicate the best average optimality for the given combination of $\delta$ and $W$. Initially, we can see that the optimality ratios tend toward 1 as $\delta$ increases, as the proportion of pairs of score widths that can be packed.

\begin{itemize}
	\item optimality converge toward 1 as $\delta$ increases, i.e. as the problem instances become less difficult.
	\item When $\delta = 1.0$. MFFD and MFFD$^+$ perform identically, as the problems are equivalent to the original BPP, and so they behave like the original FFD heuristic.
	\item When $\delta = 0.0$, all optimalities are equal, as each item requires its own strip, and so $|\mathcal{I}|$ strips are needed for all 1000 instances.
	\item Values in bold represent the best avg optimality of the three heuristics for the given level of $\delta$. 
	\item Clear pattern across both set sizes, PS performs the best out of the three for $\delta = 0.1$ to 0.7 when $W = 5000$, and for $\delta = 0.1$ to 0.6 when $W = 2500$.
	\item However, when $W = 1250$, PS is only superior when $\delta = 0.1$ and 0.2. 
	\item On further inspection, we can see that the difference between the average optimalities for PS and MFFD$^+$ is negligible.
	\item Using Table 1 as an example, when $\delta = 0.2$, the solutions generated by PS are on average within 31.781\% of $t$, and for MFFD$^+$ within 32.61\%, a difference of 0.829\%. 
	\item In practical terms, PS produces solutions using 303.95 strips on average, where $t = 230.648$. MFFD$^+$ produces solutions using 305.862 strips, a difference of less than two strips. 
	\item Another trend is that PS produces average solutions within 10\% of $t$ for $\delta = 0.5$ onwards when $W = 5000$ and 2500, however when $W = 1250$, PS only produces solutions within 10\% of $t$ for $\delta = 0.8$ onwards.
	\item In contrast, MFFD and MFFD$^+$ can produce solution within 10\% of $t$ when $W = 1250$ for $\delta = 0.5$ onwards for $n = 500$, and $\delta = 0.4$ for $n = 1000$.
	\item MFFD$^+$ can also find solutions within 1\% of $t$ for $\delta = 0.8$ onwards for all strip widths when $n = 1000$, and the two larger strip widths when $n = 500$.
	\item Number of items per strip smaller, problem is harder, falkenauer 1996, martello and toth 1990b items per strip roughly 2, triplets hardest problems
	\item all instances run in less than 160ms
	\item For each instance where opt $>1$, two cases: either $t$ is incorrect, and there does not exist a solution with $k = t$ strips, and the solution found is actually an optimal solution, or $t$ is correct, and the heuristic cannot find a solution with $t$ strips, not an optimal solution.
	\item since this is a relatively new problem, no problem instances solved to optimality, and so we have no way of knowing if in fact the heuristics are finding optimal solutions. Only able to compare with the theoretical minimum. 
	\item $W =5000$ Table 2, $\delta = 0.7$ PS has best optimality, however looking at the number of instance solved to $t$, PS has 0, whereas MFFD$^+$ has 152. Therefore there are at least 152 instances where the optimal number of strips = $t$ that PS is not able to solve to optimality.
	\item We see that PS produces solutions using fewer strips on average of the three algoritihms for a larger range of $\delta$ when the strips are wider, and MFFD$^+$ produces solutions using fewer strips for a wider range of $\delta$ when the strips are shorter.
	\item We can see that the range of values of $\delta$ in which PS produces solutions using the minimum number of strips on average of the three algorithms reduces as the strip width decreases
	\item PS never performs best for 0.8 onwards
\end{itemize}

\begin{table}[!htb]
	\centering
	\ra{1.3}
	\caption{Average optimality ratios for $n = 500$. The average number of items per strip for $W = 5000$, 2500 and 1250 are 8.4746, 4.3103, and 2.1645 respectively.}
	\begin{tabular}{@{}c@{\hspace{25pt}}c@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}c@{}}\toprule
		& \multicolumn{3}{c}{$W = 5000$, $t = 58.039$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 2500$, $t = 115.571$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 1250$, $t = 230.648$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0 & 8.61813 & 8.61813 & 8.61813 && 4.32789 & 4.32789 & 4.32789 && 2.16858 & 2.16858 & 2.16858 \\
		0.1 & 5.21367 & \textbf{4.84177} & 5.16085 && 2.65934 & \textbf{2.47735} & 2.65679 && 1.51486 & \textbf{1.47896} & 1.51486 \\
		0.2 & 4.03055 & \textbf{3.45946} & 3.97645 && 2.12099 & \textbf{1.84737} & 2.11788 && 1.32613 & \textbf{1.31781} & 1.32610 \\
		0.3 & 3.11091 & \textbf{2.34803} & 3.03816 && 1.72955 & \textbf{1.39676} & 1.70772 && 1.19451 & 1.22922 & \textbf{1.19381} \\
		0.4 & 2.43605 & \textbf{1.52907} & 2.29713 && 1.45976 & \textbf{1.12798} & 1.40951 && 1.11037 & 1.18065 & \textbf{1.10819} \\
		0.5 & 1.91057 & \textbf{1.04130} & 1.69129 && 1.26343 & \textbf{1.03341} & 1.19575 && 1.05771 & 1.15415 & \textbf{1.05319} \\
		0.6 & 1.49086 & \textbf{1.01263} & 1.24645 && 1.12448 & \textbf{1.02957} & 1.06874 && 1.02932 & 1.13470 & \textbf{1.02398} \\
		0.7 & 1.19571 & \textbf{1.01257} & 1.04468 && 1.04933 & 1.02774 & \textbf{1.01918} && 1.01741 & 1.11368 & \textbf{1.01444} \\
		0.8 & 1.04977 & 1.01203 & \textbf{1.00815} && 1.01595 & 1.02661 & \textbf{1.00790} && 1.01297 & 1.09136 & \textbf{1.01197} \\
		0.9 & 1.00929 & 1.01197 & \textbf{1.00463} && 1.00734 & 1.02586 & \textbf{1.00641} && 1.01159 & 1.07283 & \textbf{1.01139} \\
		1.0 & \textbf{1.00423} & 1.01171 & \textbf{1.00423} && \textbf{1.00626} & 1.02561 & \textbf{1.00626} && \textbf{1.01130} & 1.06549 & \textbf{1.01130} \\
		\bottomrule
	\end{tabular}	
	\label{table:n=500}
\end{table}

\begin{table}[!htb]
	\centering
	\ra{1.3}
	\caption{Average optimality ratios for $n = 1000$. The average number of items per strip for $W = 5000$, 2500 and 1250 are 8.6207, 4.329 and 2.1692 respectively.}
	\begin{tabular}{@{}c@{\hspace{25pt}}c@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}ccc@{\hspace{8pt}}c@{\hspace{8pt}}c@{}}\toprule
		& \multicolumn{3}{c}{$W = 5000$, $t = 115.534$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 2500$, $t = 230.563$} & \phantom{abcd} & \multicolumn{3}{c}{$W = 1250$, $t = 460.623$} \\
		\cmidrule{2-4} \cmidrule{6-8} \cmidrule{10-12}
		$\delta$ & MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$ && MFFD & PS & MFFD$^+$\\ \midrule
		0.0 & 8.65711 & 8.65711 & 8.65711 && 4.33803 & 4.33803 & 4.33803 && 2.17138 & 2.17138 & 2.17138 \\
		0.1 & 5.17263 & \textbf{4.84248} & 5.14027 && 2.63606 & \textbf{2.48111} & 2.64271 && 1.51056 & \textbf{1.46744} & 1.51121 \\
		0.2 & 3.97565 & \textbf{3.46163} & 3.95164 && 2.09301 & \textbf{1.85824} & 2.10161 && 1.31785 & \textbf{1.31090} & 1.31876 \\
		0.3 & 3.04866 & \textbf{2.35023} & 3.01260 && 1.69777 & \textbf{1.40862} & 1.68776 && \textbf{1.18304} & 1.22111 & 1.18344 \\
		0.4 & 2.37436 & \textbf{1.52046} & 2.26560 && 1.42646 & \textbf{1.13101} & 1.38430 && 1.09722 & 1.17063 & \textbf{1.09614} \\
		0.5 & 1.84748 & \textbf{1.02600} & 1.64169 && 1.22960 & \textbf{1.03017} & 1.16950 && 1.04411 & 1.14430 & \textbf{1.03986} \\
		0.6 & 1.43331 & \textbf{1.01212} & 1.20264 && 1.09935 & \textbf{1.02738} & 1.05010 && 1.01971 & 1.12392 & \textbf{1.01480} \\
		0.7 & 1.15500 & \textbf{1.01198} & 1.02574 && 1.03421 & 1.02569 & \textbf{1.01224} && 1.01150 & 1.10164 & \textbf{1.00926} \\
		0.8 & 1.03182 & 1.01182 & \textbf{1.00587} && 1.01111 & 1.02454 & \textbf{1.00682} && 1.00861 & 1.07791 & \textbf{1.00804} \\
		0.9 & 1.00707 & 1.01165 & \textbf{1.00419} && 1.00663 & 1.02406 & \textbf{1.00608} && 1.00784 & 1.05885 & \textbf{1.00774} \\
		1.0 & \textbf{1.00399} & 1.01169 & \textbf{1.00399} && \textbf{1.00603} & 1.02400 & \textbf{1.00603} && \textbf{1.00770} & 1.05406 & \textbf{1.00770} \\
		\bottomrule
	\end{tabular}
	\label{table:n=1000}	
\end{table}
